Privacy-Preserving Localization Using Private Linear-Combination Aggregation

### 2020-11-23 main.tex
AMEERICAAAAAAAA(N) - ugly but as requested - didn't add the last comma in lists, meh - didn't treat the word 'proof' as only a verb as US Grammarly says (still used as a noun since that makes more sense in the context)

\begin{document}
%
% paper title
\title{Privacy-Preserving Localisation Using Private Linear-Combination Aggregation}
%
%
% author names and IEEE memberships
\author{Marko~Ristic,
        Benjamin~Noack,~\IEEEmembership{Member,~IEEE,}
        and~Uwe~D.~Hanebeck,~\IEEEmembership{Fellow,~IEEE}% <-this % stops a space
\thanks{M. Ristic, B. Noack and U. D. Hanebeck are with the Intelligent Sensor-Actuator-Systems Laboratory, Institute for Anthropomatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany. e-mail: marko.ristic@kit.edu}}% <-this % stops a space
% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localisation methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localisation protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localisation filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

@@

\section{Introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localisation methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localisation methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large application due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localisation scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighbouring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localisation filter using this scheme such that sensor and navigator privacy is preserved.

The remainder of the paper is structured as follows. In Section \ref{sec:lit_overview} we give an overview of the relevant literature to the topic and in Section \ref{sec:problem_statement} we introduce the localisation problem, the relevant aggregation opertation for its solution and the restrictions on involved parties. Sections \ref{sec:crypto_prelim} and \ref{sec:lcao_definition} will summarise cryptographic preliminaries and introduce a novel security notion for our required aggregation before defining and proving an encryption scheme which satisfies it. Sections \ref{sec:loc_prelim} and \ref{sec:priv_localisation} introduce localisation preliminaries and our privacy-preserving localisation method using the private aggregation scheme. We show and discuss simulation results in Section \ref{sec:sim_and_results} and conclude our work in Section \ref{sec:conclusion}.

@@

\section{Literature Overview} \label{sec:lit_overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localisation and estimation problems.

@@

    \item[Differential Privacy] These techniques formalise preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation. To handle the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs \cite{hanPrivacyControlDynamical2018,dworkDifferentialPrivacySurvey2008,andresGeoindistinguishabilityDifferentialPrivacy2013}. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localisation.
    \item[Aggregator Oblivious Encryption] This method of encryption formalises the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

    \item[Multi-Client Function Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. They are suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalisation of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.

@@

In the context of privacy-preserving localisation and state estimation, some of the aforementioned schemes have found applications and some specialised encryption methods have been proposed. These have been summarised in the next section.

@@

\subsection{Applications to Localisation and Estimation}
Localisation methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide a computationally plausible and accurate localisation filter in the presence of range-only sensors while providing uncertainties associated with estimates.

@@

Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralised or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ however, in that weights are generalised to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants.

Along with encryption schemes and their application to localisation-relevant algorithms, encryption schemes with homomorphic properties also require suitable encoding when used with real numbers. As is the case with range sensors, real-valued measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We use a modified version of \cite{farokhiSecurePrivateControl2017} due to its simplicity and applicability, and introduce the scheme details in Section \ref{sec:loc_prelim}.

While some aspects of privacy-preserving localisation are solved in the above methods and encryption schemes, they fail to provide a localisation solution in our context. In the next section, we will formalise our problem and its accompanying security requirements.

@@

\section{Problem Statement} \label{sec:problem_statement}
The localisation scenario we consider in this work is that of Bayesian self-navigation using range-only sensors. We consider localisation in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localisation, however, may be a goal of all those involved, justifying their cooperation.

@@

\subsection{Problem Decomposition} \label{subsec:problem_decomposition}
The recursive nature of the problem definition above complicates defining and proving of security requirements for the localisation scenario. As with existing estimation methods \cite{benhamoudaNewFrameworkPrivacyPreserving2016,alexandruEncryptedCooperativeControl2019} we will reduce the communication requirements, of sensors and the navigator in our case, to an aggregation operation which will make security more easily definable.

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by a different \textit{instance} variable $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarised in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localisation} we will show how this protocol is sufficient to compute the required estimation information at the sensors needed to update the navigator's estimation filter.

@@

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled with pre-defined sensor subsets during an initial step by a trusted party. An example has been shown in Figure \ref{fig:sensor_subsets}, which reduces to two separate localisation instances with global broadcasts.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single time-step.

    In the context of wireless communication, the wide-spread use of cheap non-directional antennas make this assumption reasonable.
    \item[Honest-but-curious sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning they are assumed to follow the localisation procedure correctly but may store or use any gained sensitive information.

@@

    % Gray range circles
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \draw[gray!50] (s1) circle (\srange);
        \draw[gray!50] (s2) circle (\srange);
        \draw[gray!50] (s3) circle (\srange);
        \draw[gray!50] (s4) circle (\srange);
        \draw[gray!50] (s5) circle (\srange);
    \end{scope}
    % Red subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[red!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Blue subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \fill[blue!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Magenta double overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[magenta!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Sensors
    \fill[black] (s1) circle (\sradius);
    \fill[black] (s2) circle (\sradius);
    \fill[black] (s3) circle (\sradius);
    \fill[black] (s4) circle (\sradius);
    \fill[black] (s5) circle (\sradius);
    % Bounding box
    \draw[gray] (brect_start) rectangle (brect_end);
    % Navigator
    \clip (brect_start) rectangle (brect_end);
    \pic[rotate=-40,xscale=0.22,yscale=0.3] at (nav_loc) {plane};
    \begin{scope}[shift={(nav_loc)},rotate=-40,xscale=0.22,yscale=0.3]
        % Trail
        \fill[bottom color=white, top color=gray] (-0.05, -1.5) rectangle (0.05, -50);
    \end{scope}
    % Sensor labels
    \node[label=1] at (s1) {};
    \node[label=2] at (s2) {};
    \node[label=3] at (s3) {};
    \node[label=4] at (s4) {};
    \node[label=5] at (s5) {};
\end{tikzpicture}
\caption{An example of sensor subsets. Subsets $\{1,2,3\}$ and $\{3,4,5\}$ can be used independently for localisation when the navigator is within their ranges.}
\label{fig:sensor_subsets}
\end{figure}

@@

We formalise the security notion of Linear-Combination Aggregator Obliviousness (LCAO) as the following game between attacker and challenger:

>>>>

\begin{document}
%
% paper title
\title{Privacy-Preserving Localization Using Private Linear-Combination Aggregation}
%
%
% author names and IEEE memberships
\author{Marko~Ristic,
        Benjamin~Noack,~\IEEEmembership{Member,~IEEE,}
        and~Uwe~D.~Hanebeck,~\IEEEmembership{Fellow,~IEEE}% <-this % stops a space
\thanks{M. Ristic, B. Noack and U. D. Hanebeck are with the Intelligent Sensor-Actuator-Systems Laboratory, Institute for Anthropomatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany. e-mail: marko.ristic@kit.edu}}% <-this % stops a space
% \thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Distributed state estimation and localization methods have become increasingly popular with the rise of ubiquitous computing, and have led naturally to an increased concern regarding data and estimation privacy. Traditional distributed sensor navigation methods involve the leakage of sensor or navigator information during localization protocols, thus not preserving participants’ data privacy. Existing approaches which provide such guarantees fail to address sensor and navigator privacy in some common non-linear measurement model applications and forfeit broad applicability. We define the novel notion of linear-combination aggregation encryption and provide a cryptographically secure instance that is applied to the Extended Information Filter with range-sensor measurements. Our method keeps navigator location, sensor locations and sensor measurements private during navigation, and has been implemented to evaluate its accuracy and performance. The novel computationally plausible and provably secure range-based localization filter has direct applications to environments where participants are not trusted or data is considered sensitive.
\end{abstract}

@@

\section{Introduction}
\IEEEPARstart{L}{ocalisation} methods in distributed sensor environments have long been an active topic of research \cite{pierceIntroductionLoran1946,liContributedReviewSourcelocalization2016}. Ongoing advancements in portable computing power and sensor capabilities have led to the development of various forms of localization methods, from range-free and range-based signal strength measurements \cite{wangConvexCombinationSource2018,heRangeFreeLocalizationSchemes2003}, acoustic range estimates \cite{beutlerNewNonlinearFiltering2004} and magnetic field measurements \cite{sieblerLocalizationMagneticField2020} to the TDOA and TOA pseudo-range multilateration methods for aircraft \cite{liContributedReviewSourcelocalization2016}. Range-based localization methods, including the use of radar measurements, GPS clocks and signal strength measurements, have found particularly large applications due to their robustness and accuracy. These methods use distance measurements, or computed estimates, to produce location estimates and typically involve gathering sensor details such as measurement noise, geographical location and observations centrally. Broadcasting or gathering information in this way, however, implies a trust between participants within a trusted communication network to ensure shared data may not be used for individual malicious gain. With the advancements in cloud and ubiquitous computing, changes in possible use-cases involving distributed sensors have made the requirements of information security and user privacy more apparent than before \cite{brennerSecretProgramExecution2011,renSecurityChallengesPublic2012}. For example, an aircraft localization scenario in the presence of privately operated measurement stations may imply measurements and locations of stations should be kept private from navigating vehicles. Or in autonomous vehicle state estimation, benefit from neighboring vehicles' information may be desired while no participant wishes to share information about their hardware. Additionally, the dispatching of any expensive intermediate computations to service-providing cloud operators may require ensuring the privacy of information involved in the computations. All of these scenarios, and the one we will consider, require methods for computing specific operations while ensuring the privacy of some or all of the input data. Our contributions in this work will be presented in two parts. We first propose a novel aggregation-based encryption scheme which allows private aggregation of weighted sums, before we develop a localization filter using this scheme such that sensor and navigator privacy is preserved.

The remainder of the paper is structured as follows. In Section \ref{sec:lit_overview} we give an overview of the relevant literature to the topic and in Section \ref{sec:problem_statement} we introduce the localization problem, the relevant aggregation operation for its solution and the restrictions on involved parties. Sections \ref{sec:crypto_prelim} and \ref{sec:lcao_definition} will summarize cryptographic preliminaries and introduce a novel security notion for our required aggregation before defining and proving an encryption scheme which satisfies it. Sections \ref{sec:loc_prelim} and \ref{sec:priv_localization} introduce localization preliminaries and our privacy-preserving localization method using the private aggregation scheme. We show and discuss simulation results in Section \ref{sec:sim_and_results} and conclude our work in Section \ref{sec:conclusion}.

@@

\section{Literature Overview} \label{sec:lit_overview}
In this section, we will discuss methods for providing security guarantees in signal processing tasks, and their applications to localization and estimation problems.

@@

    \item[Differential Privacy] These techniques formalize preserving the privacy of individual datum within an aggregation operation by introducing known distributions of noise to the resulting aggregation. To handle the cases with no trusted aggregator, local differential privacy is achieved by introducing noise to individual inputs \cite{hanPrivacyControlDynamical2018,dworkDifferentialPrivacySurvey2008,andresGeoindistinguishabilityDifferentialPrivacy2013}. Although differential privacy has strong computational benefits due to its implementational simplicity, it provides relatively weak security guarantees for multi-party networks, with no guarantees in the presence of multiple colluding corrupted parties. Additionally, noisy aggregation results hinder its use in scenarios where high precision is a desired property such as aircraft localization.
    \item[Aggregator Oblivious Encryption] This method of encryption formalizes the privacy of \textit{all} aggregation data from honest parties \cite{shiPrivacyPreservingAggregationTimeSeries2011,chanPrivacyPreservingStreamAggregation2012,joyeScalableSchemePrivacyPreserving2013}. Encryptions of aggregation data are proven to be computationally indistinguishable and guarantee that only their sum can be learned by an aggregator, whether or not they are trusted. The originally proposed method \cite{shiPrivacyPreservingAggregationTimeSeries2011} is restricted to a small input domain which has been expanded and computationally simplified in later works \cite{joyeScalableSchemePrivacyPreserving2013,benhamoudaNewFrameworkPrivacyPreserving2016}. The aggregation operation in these schemes is, however, strictly limited to the sum operation and restricts its applicability to simple processing tasks such as smart-meter sums or mean calculations \cite{chanPrivacyPreservingStreamAggregation2012,benhamoudaNewFrameworkPrivacyPreserving2016}.

@@

    \item[Multi-Client Function Encryption] These schemes allow the computing of a specific function given encrypted inputs \cite{bonehFunctionalEncryptionDefinitions2011,goldwasserMultiinputFunctionalEncryption2014}. They are suitable for data-sensitive distributed processing as encryptions must be combined before a result is obtained, and can be considered a generalization of the aggregation schemes described earlier. While general schemes that allow computing arbitrary functions are computationally expensive and complicate security definitions in distributed settings \cite{agrawalFunctionalEncryptionNew2013}, partial schemes supporting only some functions, such as vector dot products \cite{chotardDecentralizedMultiClientFunctional2018}, exist with formal security proofs, but have yet to find applications in more complicated signal processing tasks.

@@

In the context of privacy-preserving localization and state estimation, some of the aforementioned schemes have found applications and some specialized encryption methods have been proposed. These have been summarized in the next section.

@@

\subsection{Applications to Localization and Estimation}
Localization methods exist for a variety of purposes requiring different levels of accuracy and supporting various forms of sensors. We aim to provide a computationally plausible and accurate localization filter in the presence of range-only sensors while providing uncertainties associated with estimates.

@@

Encrypted model-predictive-control methods \cite{farokhiPrivacyDynamicalSystems2020} also introduce some privacy-preserving methods relevant to estimation. In \cite{alexandruEncryptedCooperativeControl2019,alexandruPrivateWeightedSum2020}, the novel security notions of private weighted sum aggregation with centralized or hidden weights (pWSAc and pWSAh, respectively) are proposed. The schemes capture a weighted aggregation operation suitable for providing privacy in some control environments and are similar in principle to the aggregation decomposition proposed in this work. The requirements of our aggregation operation differ however, in that weights are generalized to arbitrary length linear-combinations, which must be consistent across all participants, and can vary over time. We also require a formal security proof capturing these weight requirements as well as the explicit capabilities of participants.

Along with encryption schemes and their application to localization-relevant algorithms, encryption schemes with homomorphic properties also require suitable encoding when used with real numbers. As is the case with range sensors, real-valued measurements need to be encoded as integers with operations consistent under scheme-provided homomorphic operations. Google's Encrypted BigQuery Client \cite{googleEncryptedbigqueryclient2015} provides an encoding which supports both addition and unencrypted multiplication when used with an additively homomorphic encryption scheme. It, however, leaks the exponent information of encrypted numbers. In \cite{farokhiSecurePrivateControl2017}, a simpler alternative is proposed, which leaks no information about encrypted numbers, but allows only a single unencrypted multiplication between encodings. We use a modified version of \cite{farokhiSecurePrivateControl2017} due to its simplicity and applicability, and introduce the scheme details in Section \ref{sec:loc_prelim}.

While some aspects of privacy-preserving localization are solved in the above methods and encryption schemes, they fail to provide a localization solution in our context. In the next section, we will formalize our problem and its accompanying security requirements.

@@

\section{Problem Statement} \label{sec:problem_statement}
The localization scenario we consider in this work is that of Bayesian self-navigation using range-only sensors. We consider localization in the two-dimensional case for simplicity but will derive methods suitable for an extension to the three-dimensional equivalent.

The navigator state is defined as 
\begin{equation}
    \rvec{x} = 
    \begin{bmatrix}
        \rv{x} & \rv{dx} & \rv{y} & \rv{dy}
    \end{bmatrix}^\top\,. \label{eqn:state_def}
\end{equation}
A known process model is followed, which at time $k$ is given by
\begin{equation}
    \rvec{x}_k = \vec{f}(\rvec{x}_{k-1})+\rvec{w}_k\,, \label{eqn:process_model}
\end{equation}
with zero-mean Gaussian process noise $\rvec{w}_k \sim \mathcal{N}(\vec{0},\mat{Q})$. The measurement model is dependant on sensor $i$ and given by
\begin{equation}
    \rv{z}_{k,i} = h_i(\rvec{x}_k)+\rv{v}_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with noise $\rv{v}_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$ at location
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
defined as
\begin{equation}
    \begin{split}
        h_i(\rvec{x}) &= \left\lVert
        \begin{bmatrix}
            \rv{x} & \rv{y}
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(\rv{x}-s_{x,i})^2 + (\rv{y}-s_{y,i})^2}\,.
    \end{split}
\end{equation}
We wish to run a state estimation filter with models \eqref{eqn:process_model} and \eqref{eqn:measurement_model} such that all involved sensors $1 \leq i \leq n$ do not learn state estimates of $\rvec{x}_{k}$ and the navigator does not learn sensor locations $\vec{s}_i,\,1 \leq i \leq n$, measurement variances $r_i$ or their measurements $\rv{z}_{k,i}$ at any time $k$. We motivate these goals with the example of aircraft navigation in the presence of privately-operated range-measuring stations. It is reasonable to assume that the current state of an aircraft may not wish to be disclosed to unknown parties and, similarly, that tower locations may wish to be kept private from unidentified navigating aircraft. The additional safety to passengers provided by accurate aircraft localization, however, may be a goal of all those involved, justifying their cooperation.

@@

\subsection{Problem Decomposition} \label{subsec:problem_decomposition}
The recursive nature of the problem definition above complicates defining and proving of security requirements for the localization scenario. As with existing estimation methods \cite{benhamoudaNewFrameworkPrivacyPreserving2016,alexandruEncryptedCooperativeControl2019} we will reduce the communication requirements, of sensors and the navigator in our case, to an aggregation operation which will make security more easily definable.

To support multiple aggregations at each time-step $k$, as is required for the aggregation of multi-dimensional data, aggregation instances have been denoted by a different \textit{instance} variable $t$. The desired bi-direction communication protocol at instance $t$ consists of of the computation of linear combinations of navigator weights by sensors, before their privacy-preserving aggregation by the navigator. That is, $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ are broadcast by the navigator, linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ are computed by each sensor $1\leq i\leq n$, and aggregation is computed back at the navigator. This has been summarized in Figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\vspace{\baselineskip}
\begin{tikzpicture}[font=\footnotesize]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and combination step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

In Section \ref{sec:priv_localization} we will show how this protocol is sufficient to compute the required estimation information at the sensors needed to update the navigator's estimation filter.

@@

    A problem caused by this requirement is apparent in the context of aircraft navigation. It may be the case that some sensors are not within the range of a navigator broadcast. We loosely propose that this scenario can be handled with pre-defined sensor subsets during an initial step by a trusted party. An example has been shown in Figure \ref{fig:sensor_subsets}, which reduces to two separate localization instances with global broadcasts.
    \item[Consistent navigator broadcast] Further, we make the assumption that broadcast information from the navigator is received equal by all sensors. This means the navigator may not send different information to individual sensors during a single time-step.

    In the context of wireless communication, the wide-spread use of cheap non-directional antennas make this assumption reasonable.
    \item[Honest-but-curious sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning they are assumed to follow the localization procedure correctly but may store or use any gained sensitive information.

@@

    % Gray range circles
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \draw[gray!50] (s1) circle (\srange);
        \draw[gray!50] (s2) circle (\srange);
        \draw[gray!50] (s3) circle (\srange);
        \draw[gray!50] (s4) circle (\srange);
        \draw[gray!50] (s5) circle (\srange);
    \end{scope}
    % Red subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[red!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Blue subgroup overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \fill[blue!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Magenta double overlap
    \begin{scope}
        \clip (brect_start) rectangle (brect_end);
        \clip (s1) circle (\srange);
        \clip (s2) circle (\srange);
        \clip (s3) circle (\srange);
        \clip (s4) circle (\srange);
        \clip (s5) circle (\srange);
        \fill[magenta!40] (brect_start) rectangle (brect_end);
    \end{scope}
    % Sensors
    \fill[black] (s1) circle (\sradius);
    \fill[black] (s2) circle (\sradius);
    \fill[black] (s3) circle (\sradius);
    \fill[black] (s4) circle (\sradius);
    \fill[black] (s5) circle (\sradius);
    % Bounding box
    \draw[gray] (brect_start) rectangle (brect_end);
    % Navigator
    \clip (brect_start) rectangle (brect_end);
    \pic[rotate=-40,xscale=0.22,yscale=0.3] at (nav_loc) {plane};
    \begin{scope}[shift={(nav_loc)},rotate=-40,xscale=0.22,yscale=0.3]
        % Trail
        \fill[bottom color=white, top color=gray] (-0.05, -1.5) rectangle (0.05, -50);
    \end{scope}
    % Sensor labels
    \node[label=1] at (s1) {};
    \node[label=2] at (s2) {};
    \node[label=3] at (s3) {};
    \node[label=4] at (s4) {};
    \node[label=5] at (s5) {};
\end{tikzpicture}
\caption{An example of sensor subsets. Subsets $\{1,2,3\}$ and $\{3,4,5\}$ can be used independently for localization when the navigator is within their ranges.}
\label{fig:sensor_subsets}
\end{figure}

@@

We formalize the security notion of Linear-Combination Aggregator Obliviousness (LCAO) as the following game between attacker and challenger:
